---
title: "Working with R in remote and virtual environments"
bibliography: [bibliography.bib, packages.bib]
biblio-style: apalike
link-citations: true
pkgdown:
  as_is: true
---

```{r, child="_common.Rmd"}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

---- 

In this Guide, we will explore two powerful strategies for working with R in remote and virtual environments: RStudio Cloud and Docker with Rocker. These tools enable you to easily collaborate on projects, manage dependencies, and ensure reproducibility of your work.

## Introduction

Working with R in remote and virtual environments can greatly enhance your productivity and collaboration capabilities. In this post, we will discuss two popular options for achieving this: RStudio Cloud and Docker with Rocker.

## RStudio Cloud

RStudio Cloud is a cloud-based platform that allows you to run RStudio in your web browser without the need for any local installation.

### What is RStudio Cloud?

RStudio Cloud provides an environment where you can create, edit, and run R projects from anywhere with internet access. It offers several advantages:

* No need to install R or RStudio locally
* Access your projects from any device
* Collaborate with others in real-time
* Easily share your work

### Setting up RStudio Cloud

To get started with RStudio Cloud, follow these steps:

1. Go to [RStudio Cloud](https://rstudio.cloud/)
2. Sign up for a free account or log in with your existing RStudio account
3. Create a new project by clicking on the "New Project" button
4. Start coding in the RStudio IDE within your browser

### Collaboration and Sharing

RStudio Cloud makes it easy to collaborate with others and share your work. You can invite collaborators to your project, allowing them to view and edit your code in real-time. To share your project, simply click on the "Share" button in the top right corner of the RStudio Cloud interface and choose the desired sharing options.

## Docker with Rocker

Docker is a platform that allows you to create, deploy, and run applications in containers. Rocker is a collection of Docker images specifically designed for running R.

### What is Docker and Rocker?

Docker enables you to package an application along with its dependencies into a container, ensuring that it will run consistently across different environments. Rocker extends this concept to R, providing pre-built Docker images with various R configurations.

Using Docker with Rocker offers several benefits:

* Reproducible environments
* Simplified dependency management
* Easy deployment and scaling

### Setting up Docker and Rocker

To start using Docker with Rocker, follow these steps:

1. Install [Docker](https://www.docker.com/get-started) on your local machine
2. Pull the desired Rocker image from [Docker Hub](https://hub.docker.com/u/rocker/)
   ```
   docker pull rocker/rstudio
   ```
3. Run a container using the pulled image
   ```
   docker run -d -p 8787:8787 -e PASSWORD=your_password --name rstudio_container rocker/rstudio
   ```
4. Access RStudio in your browser at `http://localhost:8787` and log in with username `rstudio` and the password you set

### Creating a Custom Docker Image

You can create a custom Docker image to include specific R packages or configurations. To do this, create a `Dockerfile` with the following content:

```docker
FROM rocker/rstudio

# Install additional R packages
RUN install2.r package1 package2 package3
```

Build the custom image by running:

```
docker build -t my_custom_r_image .
```

Now you can run a container using your custom image:

```
docker run -d -p 8787:8787 -e PASSWORD=your_password --name custom_rstudio_container my_custom_r_image
```

## Conclusion

In this blog post, we have discussed two powerful strategies for working with R in remote and virtual environments: RStudio Cloud and Docker with Rocker. Both options offer unique advantages and can greatly enhance your productivity and collaboration capabilities. Give them a try and see which one works best for your needs!


## References