---
title: "Choosing and setting up an IDE"
bibliography: [bibliography.bib, packages.bib]
biblio-style: apalike
link-citations: true
pkgdown:
  as_is: true
---

```{r, child="_common.Rmd"}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Overview

...

# Environment setups

## Local environments

...

## Remote environments

You can also choose to work with R in the cloud, a **remote environment**. There are a number of cloud-based options for working with R, including [RStudio Cloud](https://www.rstudio.com/products/cloud/) and [Microsoft Azure](https://learn.microsoft.com/en-us/azure/architecture/data-guide/technology-choices/r-developers-guide). These options provide a pre-configured R environment that you can access from any computer with an internet connection. The advantage of working in the cloud is that you do not need to install R or an IDE on your local computer. The disadvantage is that you will need to be connected to the internet to work with R and the free tiers for these services are limited. 

## Virtual environments

If you are new to R, you may want to consider working in the cloud to get started. If you plan to continue to work with R in the future, you will most likely want to install R and an IDE on your local computer or explore using a **virtual environment**. Virtual environments, such as [Docker](https://www.docker.com/), provide a way to use a pre-configured computing environment or create your own that you can share with others. Virtual environments are a good option if you want to ensure that everyone in your research group is working with the same computing environment. Pre-configured virtual environments exist for R through the [Rocker project](https://rocker-project.org/) and can be used locally or in the cloud. 

# References

