---
title: "Identifying data and data sources"
bibliography: [bibliography.bib, packages.bib]
biblio-style: apalike
link-citations: true
pkgdown:
  as_is: true
---

<!--  
- [ ] Revise the pinboard tags to reflect these groupings
- [ ] Check pinboard for dead links
-->

```{r, child="_common.Rmd"}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r packages, include=FALSE}
library(tidyverse) # for data manipulation
```

## Published

### Repositories 

Language-dedicated repositories are a great source of data for language research. Below I've included a listing of some of the more commonly used repositories.

```{r pinboard-repos}
read_rds("assets/listings/data_repositories.rds") |> 
  knitr::kable(caption = "Data repositories")
```

### Corpora and datasets

Below I've included a listing of corpora and datasets that are available for language research. This list is not exhaustive, but includes a few of the more common corpora and datasets used in language research.

```{r pinboard-datasets}
read_rds("assets/listings/data_corpora.rds") |> 
  knitr::kable(caption = "Corpora and language datasets")
```

### Data sharing platforms

<!--  
- [ ] Add data sharing platforms to pinboard
- [ ] Create a listing of data sharing platforms
- [ ] Include the listing here
-->

- https://dataverse.org/
- https://osf.io/
- https://www.zenodo.org/
- https://figshare.com/
- https://www.researchgate.net/
- https://www.researchsquare.com/

### Aggregated listings

The list of data available for language research is constantly growing. I've document very few of the wide variety of resources. In Table \@ref(tab:pinboard-listings) I've included attempts by others to provide a summary of the corpus data and language resources available.

```{r pinboard-listings}
read_rds("assets/listings/data_listings.rds") |> 
  knitr::kable(caption = "Aggregated listings of language corpora and datasets")
```

## Custom-built

### Application programming interfaces (APIs)

There are many APIs available for accessing language corpora and datasets. Below I've included a few of the R packages that provide access to these resources.

```{r pinboard-apis}
read_rds("assets/listings/data_apis.rds") |> 
  knitr::kable(caption = "R Package APIs to language corpora and datasets.)")
```

## Other language resources

Data for language research is not limited to (primary) text sources. Other sources may include processed data from previous research; word lists, linguistic features, *etc*.. Alone or in combination with text sources this data can be a rich and viable source of data for a research project.

```{r pinboard-other}
read_rds("assets/listings/data_experimental.rds") |> 
  knitr::kable(caption = "Other language resources")
```

