---
title: "12. Research sharing and collaboration"
pkgdown:
  as_is: true
bibliography: [bibliography.bib, packages.bib]
biblio-style: apalike
link-citations: true
---

```{r, child="_common.Rmd"}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview



## Summary

## Check your understanding

(... examples ...)

1. `r torf(TRUE)` Literate Programming, first introduced by Donald Knuth in 1984, allows the combination of computer code and text prose in one document. 
2. The programming paradigm Literate Programming is implemented through `r mcq(c(answer = "Quarto", "R", "RStudio", "GitHub"))`, a platform that facilitates the creation of a variety of output documents based on the same source code.
3. Which of the following components does a basic Quarto document *not* contain? `r mcq(c("Front-matter section", "Prose section", answer = "Back-matter section", "Code block"))`
4. To generate a PDF document in Quarto, you can edit the format attribute value in the front-matter section to `r fitb(c(answer = "pdf", "html", "word"), ignore_case = TRUE)`.
5. `r torf(TRUE)` The code block options `echo` and `include` can be used to hide the code and output, respectively.
6. `r torf(FALSE)` In Quarto, a code block, where the programming language code is entered, is bounded by three underscores (`_`).

## References