---
title: "3. Descriptive assessment of datasets"
pkgdown:
  as_is: true
bibliography: [bibliography.bib, packages.bib]
biblio-style: apalike
link-citations: true
---

```{r, child="_common.Rmd"}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

<!--  
Purpose: To explore appropriate methods for summarizing variables in datasets given the number and informational values of the variable(s).

Extend: statistic and tabular summaries

- glimpse(): dim() # nolint
- summary():  mean(), median() # nolint
- skimr::skim(): n(), mean(), median(), etc. # nolint

Extend: ggplot2 plots

- What are the main components of a plot?
  - Data
  - Aesthetics
  - Geometries
- What are other components of a plot?
  - Facets: allow for multiple plots to be displayed in a grid
    - `facet_wrap()`, `facet_grid()`
  - Scales: allow manipulation of the axes
    - `scale_x_continuous()`, `scale_y_continuous()`
  - Coordinates: allow manipulation of the coordinate system
    - `coord_flip()`, `coord_polar()`
- How can I enhace the look of my plot?
  - Themes: allow for manipulation of the look of the plot
    - `theme_bw()`, `theme_classic()`, `theme_minimal()`, `theme_void()`
  - Labels: allow for manipulation of the labels of the plot
    - `labs()`
  - Annotations: allow for manipulation of the annotations of the plot
    - `annotate()`, `geom_text()`, `geom_label()`, `geom_segment()`
-->

In this Recipe we will explore appropriate methods for summarizing variables in datasets given the number and informational values of the variable(s). This will build on our understanding of how to summarize data using statistics, tables, and plots. We will dive deeper into the use of the `skimr` package to summarize data and the `ggplot2` package to create plots. We will reinforce our understanding of the `readr` package to read in data and the `dplyr` package to manipulate data.

...


## Summary

## Check your understanding

(... examples ...)

1. `r torf(TRUE)` Literate Programming, first introduced by Donald Knuth in 1984, allows the combination of computer code and text prose in one document. 
2. The programming paradigm Literate Programming is implemented through `r mcq(c(answer = "Quarto", "R", "RStudio", "GitHub"))`, a platform that facilitates the creation of a variety of output documents based on the same source code.
3. Which of the following components does a basic Quarto document *not* contain? `r mcq(c("Front-matter section", "Prose section", answer = "Back-matter section", "Code block"))`
4. To generate a PDF document in Quarto, you can edit the format attribute value in the front-matter section to `r fitb(c(answer = "pdf", "html", "word"), ignore_case = TRUE)`.
5. `r torf(TRUE)` The code block options `echo` and `include` can be used to hide the code and output, respectively.
6. `r torf(FALSE)` In Quarto, a code block, where the programming language code is entered, is bounded by three underscores (`_`).

## References

